How to Recognize a Bad ``Bad Smell'' (using Contrast Sets)

### 2016-03-09 Manuscript.tex
Updates from ShareLaTeX

When developers cannot fully assess the value of bad smells,
it is possible to use automatic means. We assume that
developers will change code until a bad smell detector stops triggering.
That is, if the bad smell detector is {\em loc} $>$ 100, they will
redistribute the functionality of a long method into several smaller ones.
Recalling the italics in the above quote, Fowler's criteria for successful bad smell removal is the extent
to which the reorganized code has fewer errors/defects.
Data miners can be used to predict the defects in the reorganized code
so, in this paper, we will use static code defect predictors to assess
bad smells.

After making that assessment, this paper finds
that   state-of-the-art methods
(Shantnawi~\cite{Shatnawi10}  in TSE 2010;
and our own instance-based systems~\cite{me12c} from 2012) 
are less effective for bad smell assessment than a new
contrast set learning method called XTREE.  
Standard learners, such as the regression methods used by Shantnawi~\cite{Shatnawi10}
only comment on current distributions.  
  Contrast set learners   recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code reorganization. Hence, constrast set learners
are a better way to learn when code needs to be reorganized
than, say, the methods of Shantnawi~\cite{Shatnawi10}.

We recommend using XTREE as a judge to rank the value of a set
of bad smell detectors:
\begin{itemize}
    \item Consider $N$ developers each proposing a software reorganization
based on a set of bad smells $S$.
\item XTREE will (a)~learn contrast sets
from the historical record; then (b)~reject  any reorganization based on ``bad'' bad smells, i.e.
 on any member of $S_{\mathit{bad}} \in S$
that XTREE could {\em not} be found in the historical record.
\end{itemize}
Note that we    recommend  XTREE  is used to critize, but  not to create,
bad smells. 
XTREE uses frequency counts of code features seen in the historical record
to generate
predictions about the effects of changing code
attributes. Such predictions may be inaccurate due to the infamous correlation-vs-causation problem
where spurious correlations can suggest effects that are not
truly causal.
That said, if someone else's proposed bad smell detector $S_{\mathit{bad}} $ falls {\em outside} of the space of
effects found  using XTREE's analysis, then we can say that XTREE found 
{\em neither} correlation {\em or}
causal evidence that $S_{\mathit{bad}} $ can be found in the historical record.

XTREE   has certain limits. Firstly, it is an oracle that  might reject code reorganizations proposed
by developers. XTREE does not perform those refactorings.
Secondly, XTREE assumes   Fowler's bad smells
should be assessed via their contribution to software defects.   Some researchers 
 comment that   debating bad smells is  useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects. 
That said, using
defect predictors,
it is possible to operationalize XTREE
while it might be harder to otherwise operationalize the assessment of bad smells.
Hence, 
this paper assumes that {\em better} bad smells and
bettering at finding code changes that  {\em reducing} the chances
that the new code will contain defects.

The contributions of this paper are
\begin{itemize}

    \item A new kind of contrast set learner called XTREE
    that can critique plans of what to change within a data set in order
    to improve outcomes.
    \item An experimental rig for assessing the value of tools like
    XTREE that propose changes to software. That rig uses data mining
    to build an oracle for the proposed changes.
    \item Using that learner and that oracle, we offer a method for avoid code reorganizations that are irrelevant to the task of reducing the risk of introducing defects.
    \item A resolution between proponents of bad smells (e.g. Fowler, Shantnawi, Tufano~\cite{fowler99,Shatnawi10,Tufano2015}) and those who find them not useful or confusing   (e.g.Mantyla, Sj\"oberg,Yamahista et al.~\cite{Sjoberg2013,Mantyla2004,Yamashita2013}). We say
    that bad smells can guide code change, but only if those bad smells are
    first carefully assessed. 
\end{itemize}

@@

Two of the studies in \fig{smells} offers some comments on the relative importance
of the different bad smells the   Yamashita'13 study and the developer study:
    \begin{itemize}
        \item Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
        and {\em long method}.
        \item
 Note that this agreement also means that the
  Yamashita'13 study and the {\em developer survey}   concluded
        that most of the  listed code smells are {\em not} high priority issues
        requiring code reorganization. 
 \end{itemize} 
In summary, just because a developer strongly believes something
about bad smells, does
not necessary mean that that belief is useful for the current project.
Human developers can be clever, but their thinking can also be distorted
by congitive biases.
Hence it is important to double check their beliefs in, for example,
bad smells. Tools like XTREE perform that check, using models learned
from the historical record.

@@

\begin{figure*}[tbp!]
\centering
\includegraphics[width=0.8\linewidth]{figs/changes0.png}
\caption{Suggested Changes}
\label{figs:changes}
\end{figure*}

>>>>

When developers cannot fully assess the value of bad smells,
it is possible to use automatic means.

This paper proposes the XTREE contrast set learner for recognizing
``bad'' bad smells. 
Contrast set learners   recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code reorganization. Hence, constrast set learners
like XTREE can be used to  recognize ``bad''
bad smells:
\begin{itemize}
    \item Consider $N$ developers each proposing a software reorganization
based on a set of bad smells $S$.
\item XTREE will (a)~learn contrast sets
from the historical record; then (b)~reject  any reorganization based on ``bad'' bad smells, i.e.
 on any member of $S_{\mathit{bad}} \in S$
that XTREE could {\em not} be found in the historical record.
\end{itemize}
Note that we    recommend  XTREE  is used to critize, but  not to create,
bad smells. 
XTREE uses frequency counts of code features seen in the historical record
to generate
predictions about the effects of changing code
attributes. Such predictions may be inaccurate due to the infamous correlation-vs-causation problem
where spurious correlations can suggest effects that are not
truly causal.

XXX

Nevertheless, XTREE is a still a useful tool to critique and prune
irrelevant bad smells since, 
as show below, many of the bad smell detectors offered
in the literature (and supported by readily available tools)
fall outside of  space of
effects found  using XTREE's analysis.
That is,   XTREE can find 
{\em neither} correlation {\em nor}
causal evidence that those bad smells
have any support   in the historical record.

XXXX

The same cannot be said for other methods that try to learn
bad smell thresholds. \tion{outs}
discusses {\em outlier} methods and {\em centroid delta} methods
for learning bad smell thresholds.
The {\em  outlier methods}  used by
Shatnawi, Alves and Herman et al.~\cite{Shatnawi10,Alves2010,hermans15}
propose thresholds on all code metrics with outlier values.
Similarly, the older {\em centroid delta} methods  used in our 
prior work~\cite{me12c} also returns thresholds on most of the code
metrics. That is, outlier methods and centroid deltas methods
seem to be saying that many bad smell thresholds are possible.

we proposed CD, short for
``Centroid Deltas'', as a way to find bad smell thresholds:
\begin{itemize}
    \item Clustered
data and generated centroids for each cluster;
\item Computed contrast sets (i.e. the bad smells)
as the delta between attributes in adjacent centroids $C_1 - C_2$
where examples in $C_1$ had more defects than $C_2$.
\end{itemize}
As shown below, Centroid Deltas
generates change
recommendations that are far less effective than XTREE.
Worse still Centroid Deltas reported bad smells
for far more attributes than XTREE. Hence, that prior work
is less useful than XTREE at judging the  bad smells proposed by developers.

XTREE   has certain limits. Firstly, it is an oracle that  might reject code reorganizations proposed
by developers. XTREE does not perform those refactorings.
Secondly, XTREE assumes   Fowler's bad smells
should be assessed via their contribution to software defects
(evidence: see the section in italics of the quote
that starts this introduction). Some researchers 
 comment that   debating bad smells is  useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects. 
That said, using
defect predictors,
it is possible to operationalize XTREE
while it might be harder to otherwise operationalize the assessment of bad smells.
Hence, 
this paper assumes that {\em better} bad smells and
bettering at finding code changes that  {\em reducing} the chances
that the new code will contain defects.

\newpage
The contributions of this paper are
\begin{itemize}

    \item A new kind of contrast set learner called XTREE
    that   critique plans of what to change   in order
    to improve outcomes.
    \item An experimental rig for assessing the value of tools like
    XTREE that propose changes to software. That rig uses data mining
    to build an oracle for the proposed changes.
    \item Using that learner and that oracle,
    a demonstration that XTREE is better than our prior approach~\cite{me12c}
    as well as   state of the art methods~\cite{Shatnawi10,Alves2010,hermans15}
    at avoiding code reorganizations that are irrelevant to the task of reducing the risk of introducing defects.
    \end{itemize}
    Based on the above, we can offer
    a resolution between proponents of bad smells (e.g. Fowler, Shantnawi, Tufano~\cite{fowler99,Shatnawi10,Tufano2015}) and those who find them not useful or confusing   (e.g.Mantyla, Sj\"oberg,Yamahista et al.~\cite{Sjoberg2013,Mantyla2004,Yamashita2013}). We say
    that bad smells can guide code change, but only if those bad smells are
    first carefully assessed.

@@

Two of the studies in \fig{smells} offers some comments on the relative importance
of the different bad smells the   Yamashita'13 study and the developer study:
    \begin{itemize}
        \item Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
        and {\em long method}.
        \item
 Note that this agreement also means that the
  Yamashita'13 study and the {\em developer survey}   concluded
        that most of the  listed code smells are {\em not} high priority issues
        requiring code reorganization. 
 \end{itemize} 
In summary, just because a developer strongly believes something
about bad smells, does
not necessary mean that that belief is useful for the current project.
Human developers can be clever, but their thinking can also be distorted
by cognitive biases.
Hence, as shown in \fig{smells}, developers and text books and tools 
can disagree on which bad smells are important.
Hence it is important to have some automated tool
that can assess their beliefs in, for example,
bad smells.

\section{Learning Bad Smell Thresholds}

After a review of the literature, we reached the same conclusion
as Shatnawi~\cite{Shatnawi10}; i.e. there is a dearth of papers recommending specific thresholds
for bad smells. In the recent literature (since 2010), we could
only find the handful   papers discussed below.

In the recent literature, there are two approaches for
learning bad smell thresholds:
\begin{itemize}
    \item Approaches based on {\em outlier statistics}~\cite{erni96,bender99}
    as used by Shatnawi~\cite{shatnawi10}, Alvenes et al.~\cite{Alves2010}
    and Hermans et al.~\cite{hermans15};
    \item  Approaches based on {\em cluster deltas} that we developed
    for   Centroid Deltas~\cite{me12c} and XTREE.
\end{itemize}
The outlier approach assume that unusually large measurements are risk-prone.
Hence, they generate one bad smell threshold for any metric
with such an unusually large measurements.

The advantage of outlier
methods is that they are simple to implement. However,
their disadvantage are two-fold. 
Firstly, the outlier approach is {\em verbose}; i.e.  it can
propose thresholds on most of the code metrics.
Such verbosity makes it hard   to critique and prune away proposed code reorganizations since any bad smell mentioned by a developer has some support
somewhere amongst all those proposed thresholds.

Secondly, they do not reflect
on sets of code metrics that change together. For example,
one way to remove unusually large methods is to reorganize that method's
functionality into several smaller ones. If that is done,
this means that {\em reducing} lines of code in one method
will {\em increase} the coupling of the methods of that class to
either itself or to other classes (depending on whether or not
the new smaller methods are placed in the original class, or in other classes).
Hence, it can be counter-productive to act on bad smell thresholds for {\em both}
large lines of code {\em and} seemingly excessive coupling.

Borges and Menzies first proposed the centroid delta approach to address
this second problem. Centroid deltas assume...

One is {\em succinct}; i.e. comments
on too many code metrics. If the

class' coupling mea
\begin{itemize}
    \item 
\end{itemize}that they do not consider the associations
between variables.

Cluster deltas methods assume that at
Prior work on deriving metric thresholds from benchmark data
used {\em outlier} statistics to identify abnormally large ranges. 
Another approach, used by

For example,
one of the first methods of finding thresholds was proposed Erni and Lewerentz~\cite{erni96}. Their technique to identify thresholds was based on the data distribution, specifically the mean and the standard deviation of the metric values. They propose the use of values that lie beyond one standard deviation from the mean as a threshold. The minimum value $T_{min}$ is given by $T_{min}=\mu-\sigma$, and this is used when metric definition considers very small values as an indicator of problems. Otherwise, $T_{max}=\mu+\sigma$ is used, when large metric values are considered problematic.

Several researchers~\cite{shatnawi10}~cite{alves10} have pointed out that this method is subject to a few problems. Firstly, it doesn't consider the fault-proneness of classes when the thresholds are computed. Secondly, there is a lack of empirical validation of this methodology, which impedes reasonable comparisons.

@@

\begin{figure*}[!y]
\centering
\includegraphics[width=0.90\linewidth]{figs/changes0.png}
\caption{Suggested Changes}
\label{figs:changes}
\end{figure*}
